Lattice Boltzmann study of mass transfer for Bretherton/Taylor bubble train flow

### 2012-04-24 Text/introduction.tex
Finished till validation part.

Three possible scenarios of conducting numerical simulations are examined in this work: 
\begin{enumerate}
\item The bubble train multiphase simulations are conducted by applying periodic boundary
conditions \cite{kuzmin-binary2d,kuzmin-binary3d}. 
%One
%can make an assumption of same boundary
%conditions applied to the tracer concentration to simulate mass transfer. 
Therefore, only one unit
cell is simulated with
periodic boundary conditions, see Fig. \ref{fig:benchmark}. As well, no tracer leaves a
domain. The situation resembles the plug flow situation, where no tracer can leave a unit
domain. From the physical point of view this case resembles most the mass transfer for small Capillary numbers
$Ca<0.1$.  The liquid slug is well mixed and one have a homogenuous concentration in the unit
cell, i.e. the outlet and the inlet concentrations are close. Though easier to implement, it rises certain criticism
about the inlet concentration to be equal to the outlet concentration. In the continuous
formulation as it is seen in experiments there is the concentration difference between inlet and
outlet involving the volumetric
mass transfer coefficient, see second equation in the system \ref{theor:one:concentration:time}.

One needs also address the choice of the characteristic concentration. For example, the
characteristic concentration $\langle C(t) \rangle$ required for calculation the volumetric mass
transfer coefficient, Eq.
\ref{theor:one:concentration:time}, can be taken as the average concentration in the domain:
\beq
\label{average:concentration:definition}
C(t)=\frac{\int_{liquid}{C \mathrm{d}V}}{\int{\mathrm{d}V}}.
\feq
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/benchmark_middle.eps}\\
\caption{The two-dimensional benchmarks for the  the mass transfer
coefficient (bottom) for the bubble located near the entrance (top) and at the middle of the domain
(bottom). \label{fig:benchmark}}
\end{figure}
\item The second approach is the same as the first one but the characteristic concentration is taken
as in
\cite{vanbaten-circular}. The characteristic concentration is taken as the inlet/outlet
concentration. For the two-dimensional flow between plates inlet/outlet concentrations can be found
as:
\beqal
&\langle \cinlet(t) \rangle=\frac{\int{U(y) C(0,y,t) \mathrm{d}y}}{\int{U(0,y) \mathrm{d}y}}\\
&\langle \coutlet(t) \rangle=\frac{\int{U(y) C(\lunit,y,t) \mathrm{d}y}}{\int{U(\lunit,y)
\mathrm{d}y}}\\
&\cinlet(\bm{x},t)=\coutlet(\bm{x},t),\,\text{due to periodicity}.
\feqal
Therefore, the assumptions  of this approach is
that the difference between inlet and outlet is not
considerably large and the tracer is well mixed in the slug, so that the inlet or outlet
concentrations equal to the average concentration.

\item
The approach of \citet{vanbaten-circular} is examined. Periodic boundary conditions were used. We
used the
definition of the mass transfer coefficient to be proportional to the gain of the mass in the system
divided by the concentration difference
multiplied on the surface area:
\beq
\vol=\frac{\dot{m}}{P \Delta C}\frac{P}{V}=\frac{\dot{m}}{V (\cstar-\langle C(t) \rangle)},
\feq
where mass difference in the domain can be calculated as:
\beq
\dot{m}=\frac{m_2-m_1}{t_2-t_1}=\frac{\int_{liq}{C(\bm{x},t_2)\mathrm{d}V}-\int_{liq}{
C(\bm{x},t_1)\mathrm { d } V} } {
t_2-t_1 },
\feq
where the integration element $\mathrm{d}V=\mathrm{d}x\mathrm{d}y=\mathrm{d}\bm{x}$.
In the approach of \citet{vanbaten-circular} the inlet and outlet concentrations were taken as the
characteristic concentration $\langle C(t) \rangle $.

%\item 
%The approach of the opened boundaries was examined, see Fig. \ref{fig:benchmark:jos}. At the inlet
%and outlet boundaries the diffusion flux was imposed to be zero, $\frac{\partial C}{\partial x}=0$.
%The characteristic concentration was taken as the average concentration in the whole domain.
%\begin{figure}
%\includegraphics[width=\textwidth]{Figures/benchmark_jos.eps}
%\caption{The numerical mass transfer benchmark with open boundaries. The mass transfer is
%calculated according to Eq. \ref{theor:average:concentration:time} with the characteristic
%concentration accoring to Eq. \ref{average:concentration:definition}. \label{fig:benchmark:jos}}
%\end{figure}
\item 
The last approach to be examined in this paper is the approach of the simulation of a lot of unit
volumes, see Fig. \ref{fig:benchmark:alot}. This situation corresponds to the simulation of the
bubble train head, after injecting to the pipe and travelling along the pipe. One can see that this
situation corresponds most to the experimental picture. By simulation of a certain number of the
bubble train head, the influence of the boundaries can be
diminished. In this case, there is no ambiguity in the choice of the
characteristic concentration in this case. The average concentration in time will be governed
by Eq.
\ref{theor:continuous:mass:transfer}. This simulation statement is of most correspondence to
experiments and should give right answer. However, one of the
disadvantages is to simulate a certain number of unit cells ($1$ to $10$ in this case), thus
increasing a computational
complexity.
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/benchmark_alot.eps}\\
\caption{Benchmark for a lot of units. \label{fig:benchmark:alot}}
\end{figure}
\end{enumerate}

One can notice that all the examined cases are extreme cases of one equation. The choice of boundary
conditions to use  or how many unit cells to simulate are not of importance but one needs only to
understand a proper mass balance for medium, which states:
\beq
\label{main:main:main}
\vol=\frac{\dot{m}-\int{\coutlet(t) u(\lunit,y)\mathrm{d}y}+\int{\cinlet(t) u(0,y)\mathrm{d}y}}{V
\Delta C}, 
\feq
where $\Delta C=\cstar-\langle C(t) \rangle$ with $\langle C(t) \rangle$ to be the average
concentration in whole liquid domain. Eq. \ref{main:main:main} is nothing else but a proper mass
balance: whatever was generated by the bubble surface equals to the mass domain change minus
whatever left domain plus whatever entered domain. Periodic boundary condition is the extreme
limiting case of Eq. \ref{main:main:main} as $\int{\coutlet(t)
u(\lunit,y)\mathrm{d}y}=\int{\cinlet(t) u(0,y)\mathrm{d}y}$. Thus, the mass change in the domain is
only because of the generating bubble surface. In this case, to calculate the volumetric mass
transfer coefficient one needs to track not only spatial information to calculate the averaged over
domain characteristic concentration  $\langle C(t) \rangle$ but as well time information to
calculate the mass change in the domain. It will be shown later, that for $Ca>0.7$ where velocity
pattern is changed, one can use only spatial information to calculate $\int{\cinlet(t) u(0,y)
\mathrm{d}y}$ and $\int{\coutlet(t) u(\lunit,y)\mathrm{d}y}$ as far as mass change in the domain is
zero, i.e. $\dot{m}=0$.

>>>>

Four possible scenarios of  numerical simulations are examined in this work: 
\begin{enumerate}
\item % The bubble train multiphase simulations are conducted by applying periodic boundary
%conditions \cite{kuzmin-binary2d,kuzmin-binary3d}. 
%One
%can make an assumption of same boundary
%conditions applied to the tracer concentration to simulate mass transfer. 
One unit
cell is simulated with
periodic boundary conditions, see Fig. \ref{fig:benchmark}. 
Thus, no tracer leaves a
domain like for the plug flow. 
%From the physical point of view this case resembles the mass transfer for small Capillary numbers
%$Ca<0.1$.  The liquid slug is well mixed and one have a homogeneous concentration in the unit
%cell, i.e. the outlet and the inlet concentrations are close. 
Though easier to implement, it rises certain criticism
about the inlet concentration to be equal to the outlet concentration. As it was discussed, in experiments there is the concentration difference between the inlet and
 the outlet.

The volumetric mass transfer coefficient is calculated by Eq. \ref{theor:one:concentration:time}.
The
characteristic concentration $\langle C(t) \rangle$ required for the volumetric mass
transfer coefficient, Eq.
\ref{theor:one:concentration:time}, is taken as the average concentration in the domain:
\beq
\label{average:concentration:definition}
C(t)=\frac{\int_{liquid}{C \mathrm{d}V}}{\int{\mathrm{d}V}}.
\feq
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/benchmark_middle.eps}\\
\caption{The two-dimensional benchmarks for the  the mass transfer
coefficient (bottom) for the bubble located near the entrance (top) and at the middle of the domain
(bottom). \label{fig:benchmark}}
\end{figure}
\item The second approach is to apply periodic boundary conditions as in the first case but the characteristic concentration is taken as the inlet/outlet
concentration \cite{vanbaten-circular}:
\beqal
&\langle \cinlet(t) \rangle=\frac{\int{U(y) C(0,y,t) \mathrm{d}y}}{\int{U(0,y) \mathrm{d}y}}\\
&\langle \coutlet(t) \rangle=\frac{\int{U(y) C(\lunit,y,t) \mathrm{d}y}}{\int{U(\lunit,y)
\mathrm{d}y}}\\
&\cinlet(\bm{x},t)=\coutlet(\bm{x},t),\,\text{due to periodicity}.
\feqal
Therefore, the assumptions  of this approach is
that the difference between inlet and outlet is not
considerably large and the tracer is well mixed in the slug. Thus, the inlet or outlet
concentrations equal to the average concentration.

\item
The approach of \citet{vanbaten-circular} is examined. Periodic boundary conditions were used. The work \cite{vanbaten-circular}
calculates the mass transfer coefficient as the gain of the mass in the system
divided by the concentration difference
multiplied on the surface area:
\beq
\vol=\frac{\dot{m}}{P \Delta C}\frac{P}{V}=\frac{\dot{m}}{V (\cstar-\langle C(t) \rangle)},
\feq
where mass difference in the domain can be calculated as:
\beq
\dot{m}=\frac{m_2-m_1}{t_2-t_1}=\frac{\int_{liq}{C(\bm{x},t_2)\mathrm{d}V}-\int_{liq}{
C(\bm{x},t_1)\mathrm { d } \bm{x}} } {
t_2-t_1 }.
\feq
In the approach of \citeauthor{vanbaten-circular} the inlet and outlet concentrations were taken as the
characteristic concentration $\langle C(t) \rangle $.

%\item 
%The approach of the opened boundaries was examined, see Fig. \ref{fig:benchmark:jos}. At the inlet
%and outlet boundaries the diffusion flux was imposed to be zero, $\frac{\partial C}{\partial x}=0$.
%The characteristic concentration was taken as the average concentration in the whole domain.
%\begin{figure}
%\includegraphics[width=\textwidth]{Figures/benchmark_jos.eps}
%\caption{The numerical mass transfer benchmark with open boundaries. The mass transfer is
%calculated according to Eq. \ref{theor:average:concentration:time} with the characteristic
%concentration accoring to Eq. \ref{average:concentration:definition}. \label{fig:benchmark:jos}}
%\end{figure}
\item 
The last approach to be examined in this paper is the simulation of a few unit cells, see Fig. \ref{fig:benchmark:alot}. This situation corresponds to the simulation of the
bubble train head, after injecting to the pipe and travelling along the pipe. One can see that this
situation corresponds most to the experimental picture. By simulation of a certain number of the
bubble train head, the influence of the boundaries can be
diminished. In this case, there is no ambiguity in the choice of the
characteristic concentration. The average concentration of any unit cell far away from boundaries will be governed
by Eq.
\ref{theor:continuous:mass:transfer}. This simulation statement is of most correspondence to
experiments and should give right answer. However, one of the
disadvantages is to simulate a certain number of unit cells ($1$ to $10$ in this case), thus
increasing a computational
burden.
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/benchmark_alot.eps}\\
\caption{Benchmark for a lot of units. \label{fig:benchmark:alot}}
\end{figure}
\end{enumerate}

One can notice that all examined cases are extreme cases of one equation:
\beq
\label{main:main:main}
\vol=\frac{\dot{m}-\int{\coutlet(t) u(\lunit,y)\mathrm{d}y}+\int{\cinlet(t) u(0,y)\mathrm{d}y}}{V
\Delta C}, 
\feq
where $\Delta C=\cstar-\langle C(t) \rangle$ with $\langle C(t) \rangle$ to be the average
concentration in whole liquid domain, $\dot{m}$ is the mass gain in the domain,$\int{\cinlet u(0,y) \mathrm{d}y}$ and $\int{\coutlet u(\lunit,y)\mathrm{d}y}$ are inlet/outlet mass fluxes. Eq. \ref{main:main:main} describes the mass
balance: whatever was generated by the bubble surface equals to the mass domain change minus
whatever left domain plus whatever entered domain.

Periodic boundary condition is the extreme
limiting case of Eq. \ref{main:main:main}: 
\beqstar
\int{\coutlet(t)
u(\lunit,y)\mathrm{d}y}=\int{\cinlet(t) u(0,y)\mathrm{d}y}. 
\feqstar
Then the mass change in the domain is
only because of the generating bubble surface. However, to calculate the mass transfer coefficient one needs to have characteristic concentration  $\langle C(t) \rangle$ and mass gain $\dot{m}$ which do require the storage of time and space information.

Another limiting case (will be shown later) is hydrodynamic pattern change $Ca>0.7$. The mass accumulation rate equals zero in this case, $\dot{m}=0$. One can use only spatial information to calculate the mass transfer coefficient, i.e. $\int{\cinlet(t) u(0,y)
\mathrm{d}y}$ and $\int{\coutlet(t) u(\lunit,y)\mathrm{d}y}$ require only space information.
